---
title: "MThacker_UKCars"
author: "Matthew Thacker"
date: "November 24, 2016"
output: html_document
---
#UK Cars Assignment for Unit 11

###Setup
```{r, warning=FALSE, message=FALSE}
library(fpp)
library(ggplot2)
data(ukcars)
```

There are definitive seasonal flucuations. There appears to be a general upward trend in car sales over time.

```{r}
plot(ukcars)
```

These result support the conclusion that there is seasonal variation and an upward trend in cars sales.

```{r}
fitd <- decompose(ukcars)
plot(fitd)
eeadj <- seasadj(fitd)
plot(eeadj)
```

When viewing the seasonal adjusted data for the outlier, I didn't see a significant impact beyond the quarter where the outlier was inserted.

```{r}
ukcars2 <- ukcars
ukcars2 <- ts(c(ukcars[1:54],ukcars[55]+500,ukcars[56:113]),start=c(1977,1),frequency=4)
plot(ukcars2)
fitd2 <- decompose(ukcars2)
eeadj2 <- seasadj(fitd2)
plot(eeadj2)
```

Once again the outlier seems to have only impacted the quarter where it was placed. 

```{r}
ukcars3 <- ukcars
ukcars3 <- ts(c(ukcars[1:99],ukcars[100]+500,ukcars[101:113]),start=c(1977,1),frequency=4)
plot(ukcars3)
fitd3 <- decompose(ukcars3)
eeadj3 <- seasadj(fitd3)
plot(eeadj3)
```

```{r}
decom <- stl(ukcars, s.window=5)
plot(decom)

plot(ukcars, col="gray", main="UK car sales", ylab="Sales", xlab="Time")
lines(decom$time.series[,2], col="red", ylab="Trend")
```